---
title: 'Statistical report: Digoxin in HFrEF'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
---

\newpage 
\tableofcontents 
\listoffigures
\listoftables
\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB3, https://kiheartfailure.github.io/shfdb3/, v 3.1.5. 

## Inclusion/exclusion criteria

```{r flow}
mykable(flow, caption = "Flowchart")
```

*if a patient dies immediately after the index visit he/she will not 
collect a drug prescribed at the index visit and thereby be defined as a digoxin user. 
Throughout, the index date is moved 14 days forward (so the time to outcome is started at 0). 

First patient in: `r min(pdata$shf_indexdtm)` and last patient in: `r max(pdata$shf_indexdtm)`.  

The median age (IQR) is `r pdata %>% summarise(med = dF(median(shf_age), dig = 1),
                                             q1 = dF(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = dF(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r pdata %>% count(shf_sex) %>%
  mutate(perc = dF(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    
  
## Created variables 

```{r atc}
mykable(metalm, caption = "Treatments from DDR")
```

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Missing data was imputed with multiple imputation using mice [@mice] for 10 
datasets and 10 iterations. Variables included in the model are indicated in 
Table \ref{tab:tab1}. The primary outcome, first heart failure hospitalization/death,
was included as the Nelson-Aalen estimator. Digoxin use is not included in the imputation model.  

## Propensity scores

A propensity score for treatment with digoxin was estimated for each patient with 
logistic regression for each of the 10 imputed datasets using the variables 
indicated in Table \ref{tab:tab1}. 1:1 matching without replacement [@match] 
was thereafter performed on the average of the resulting 10 ps [@psmatch]. Matching was 
allowed if the propensity score differed by 0.01 or less. The ability of the ps 
matching to balance the baseline characteristics was assessed by 
standardized mean differences.  

1:1 has deemed the best option when taking into account the number of patients 
retained and sample size. Other matching ratios: `r matchingn`. 

## Digoxin over time

The denominator is the population in the analysis (N = `r nrow(pdata)`). 
The numerator consists of the number of patients that have at least one digoxin 
prescription. The calculation is done per year, so a patient without digoxin 
at index in SwedeHF, for example 2017, might still be included in the numerator for example for 2013 or 2018.  

```{r, child = "../src/overtime.Rmd"}

```

## Baseline characteristics

```{r, child = "../src/tab1.Rmd"}

```

\clearpage
\newpage

## Associations with digoxin use

Odds ratios are estimated using logistic regression with digoxin use as outcome. 

```{r, child = "../src/ortab.Rmd"}

```

### Assumptions

Outliers were investigated with Cook's distance and multicollinearity 
with the variance inflation factor in the full model in the overall population.
=> No action deemed necessary. 

\clearpage
\newpage

## Association digoxin vs outcomes

The following outcomes are considered: 

- All-cause mortality/first HF hospitalization (primary endpoint)
- All-cause mortality
- First HF hospitalization

Data were censored at 2018-12-31 or death/emigration, 
so this assumes that the risk of hf hospitalization 
would have been the same for a censored patient had the censoring not occurred. 

Incidence per 1000 py was calculated with 95% Poisson confidence intervals. 

The outcomes are presented with the 1 - Kaplan-Meier curves
and tested with the log-rank test for the whole and matched population. 
Note that the log-rank test is theoretically 
redundant since the cox regression gives the same information. 
Note further that the cox regression for the matched population incorporates 
a frailty term taking into account the dependency between pairs 
which is not done in the log-rank test. 

Cox proportional hazards regressions were used to evaluate the association 
between digoxin and the respective outcomes. 
In the matched cohort the matched pairs were modelled using a frailty term. 

### Consistency analysis

#### Crossover

In the main analysis digoxin use is considered according to the ITT principle 
(digoxin at baseline). As a consistency analysis digoxin was included as a 
time-dependent variable in the model. 

In patients treated with digoxin at baseline the patient was assumed to be on digoxin 
until LAST dispension + 3 months or end of follow-up 
(death/emigration/2018-08-01/event) - 5 months independent of how 
long time between dispensions. In untreated patients, crossover was defined 
as a new digoxin dispension in the Dispensed Drug Registry and 
thereafter in the same manner as for patients treated with digoxin at baseline. 
That means that patients with digoxin at baseline were allowed to cross 
over max once (Yes-No) and patients without digoxin at baseline were 
allowed to cross over max twice (No-Yes-No). 

#### Competing risk

The analyses were also performed using a sub-distributional hazards model [@fg] 
were death was treated as a competing event. 

```{r, child = "../src/km.Rmd"}

```

```{r, child = "../src/outtab.Rmd"}

```

### Subgroups

Performed using interaction effects on the patients in the matched cohort and 
presented in a Forest plot. 
Note that there is missing data for some of the variables so the matching 
(and thereby the adjustment) may therefore not be complete. 

```{r, child = "../src/outtabsub.Rmd"}

```

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the primary outcome 
in the ps matched cohort (digoxin according to ITT). 
Possible outliers were visually inspected by plotting the dfbetas. 
=> `r paste0(stratavars, collapse = ", ")` were deemed not to have proportional
hazards and therefore stratified for in the cox regressions.  

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/digoxinhfref. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
